---
title: 'Replicating a Study that Analysed Fossil Fuel Tax Levels in 2003-2015'
author: "Mark Roche"
date: "11/04/22"
output: pdf_document
---

# 1) INTRODUCTION 

# Outline of Original Paper
  - Mahdavi et al (2020) ‘Why Do Governments Tax or Subsidise Fossil Fuels?’

  - Measured fossil fuel tax and subsidy levels and change between 2003-2015 and analysed     whether either varies under different political and economic settings
  
  - Relationship between fossil fuel tax/subsidies and political and economic factors

# Data: Outcome and Predictor Variables
  
  - Outcome variable: Implicit Fossil fuel tax/subsidy level; measured by comparing a        country's fuel and gas prices against the international supply price of fuel. 
 
  - Authors argue that if the local price is above the international supply price, it        implies that the country is likely a net taxer of oil and gas and vice-versa for         subsidisers. 
  
  - In other words, it is the difference between the price of fuel and the price it is to     bring the fuel to consumers, implies that a subsidy has been put in place by             government
  
  - Economic explanatory variables: GNI per-capita, government debt and fossil fuel         wealth. 
  
  - Political explanatory variables: Democracy and Government Effectiveness 


# Testing Findings of Original Paper
  - I analyse  explanations for  fossil fuel tax/subsidy levels in 2003-2015 via             cross-national data
  
  - I Do not analyse reasons for fossil fuel tax/subsidy change from 2003-15 (i.e. across time)
  
  - This tests two out of four major findings of original paper:
    - Finding 1: The relationship between tax/subsidies and economic predictors is by far                 the strongest. 
    - Finding 2: There is a weak relationship between tax/subsidies and political                         predictors. 
    
# Adding an Interaction Term 
  - Particularly wishes to expand on original study by focusing on the European Union and     whether a member state's fuel wealth explained its 2003-2015 tax and subsidy levels
  
  - This is achieved by adding an interation term between the EU countries and fuel         wealth indicators 


# Method 

# 1) Data Collection: 
  - Collected the Data From the Harvard Database Website
  - Load Data into RStudio 
  
# 2) Data Analysis I: Analyse Expanations for Tax and Subsidy Levels
  - Run a glm() linear regression to regress tax/subsidy levels on economic and political     predictors.
  - Linear regression is used as the outcome is continuous
  
     OVERALL, RESULTS FOUND WERE QUITE SIMILAR TO ORIGINAL PAPER

# 3. Data Analysis II: Analyse Whether EU Country's Wealthy From Oil and Gas Tax and         Subsidise Less
  - Run a glm() linear regression to regress tax/subsidy change on economic and political     predictors.
  - Linear regression used as outcome variable is continuous



# 2) ANALYSIS PART I: EXPLAINING  TAX AND SUBSIDY LEVELS  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("/Users/user/Desktop/Semester 2/Applied Stats II/Replication Project/Code")

```

```{r}
# Libraries
library(dplyr)
library(tidyr)
library(ggplot2)

# Load dataset
dataset <- read.csv('cross_national_ffs_final.csv')
View(dataset)

### 1) RUN REGRESSION MODEL ###
## Try with glm()
lm <- glm(meanbmgap2015adj ~ meanfuelexports+fuel_income_dependence+average_oilgas_exports_pc + 
           meangdppc + meangdppcatlas   + gdp_ppc_gd    +  autocracy_polity  + meangoveffect +
           avg_gov_debt  +  meanvat, data = dataset, family=gaussian(link="identity"))

# Summarise the output
summary(lm)
```

####### DATA VISUALISATION AND INTERPRETATION ####### 

With the explanatory and outcome variable in place in a summarised regression, the covariates can now be interpreted and visualised using ggplot. 

This can test whether findings observed in the original paper are valid. 


# ECONOMIC PREDICTORS:

# Fossil Fuel Wealth: Fuel income dependence, meanfuelexports and average oil exports per capita. 

# Fuel income dependence
  - Strongly negative correlation between fuel income dependence and net implicit gas        tax/subs
```{r}
ggplot(dataset, aes(fuel_income_dependence, meanbmgap2015adj)) +           
  geom_point() +                                     
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth")
```

# Average Fuel Exports 
```{r}
ggplot(dataset, aes(log(meanfuelexports), meanbmgap2015adj)) +           
  geom_point() +                                     
  stat_smooth(method = "lm",
              formula = y ~ x + I(x^2),
              geom = "smooth")

```

# Average Oil Exports Per Capita
```{r}
ggplot(dataset, aes(average_oilgas_exports_pc, meanbmgap2015adj)) +           
  geom_point() +                                     
  stat_smooth(method = "lm",
              formula = y ~ x + I(x^2),
              geom = "smooth")
```
# GNI Per Capita 
```{r}
ggplot(dataset, aes(log(meangdppcatlas), meanbmgap2015adj)) +           
  geom_point() +                                     
  stat_smooth(method = "lm",
              formula = y ~ x + I(x^2),
              geom = "smooth")
```
# Average Government Debt
```{r}
ggplot(dataset, aes(log(avg_gov_debt), meanbmgap2015adj)) +           
  geom_point() +                                     
  stat_smooth(method = "lm",
              formula = y ~ x + I(x^2),
              geom = "smooth")
```


# POLITICAL PREDICTORS

# Democracy
```{r}
ggplot(dataset, mapping = aes(autocracy_polity, meanbmgap2015adj)) +           
  geom_point() +                                     
  stat_smooth(method = "lm",
              formula = y ~ x + I(x^2),
              geom = "smooth")
```

# Government Effectiveness
```{r}
ggplot(dataset, mapping=aes(meangoveffect, meanbmgap2015adj)) +           
  geom_point() +                                     
  stat_smooth(method = "lm",
              formula = y ~ x + I(x^2),
              geom = "smooth")
```

# INTERACTION ANALYSIS: HOW DOES THE EUROPEAN UNION VARIABLE INTERACT WITH FOSSIL FUEL WEALTH


```{r}
imodel1 <- glm(meanbmgap2015adj ~ fuel_income_dependence*europeanunion, data = dataset, family=gaussian(link="identity"))

# Summarise the output
summary(imodel1)

ggplot(dataset, mapping=aes(fuel_income_dependence*europeanunion, meanbmgap2015adj)) +  
  geom_point() +                                     
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth")

# Analyse interaction between avg oil exports per capita and EU members
imodel2 <- glm(meanbmgap2015adj ~ average_oilgas_exports_pc*europeanunion, data = dataset, family=gaussian(link="identity"))

summary(imodel2)

ggplot(dataset, mapping=aes(meanfuelexports*europeanunion, meanbmgap2015adj)) +
  geom_point() +                                     
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth")

interaction.plot(
x.factor = dataset$fuel_income_dependence,
trace.factor = dataset$europeanunion,
response = dataset$europeanunion,
fun = median,
ylab = "Implicity Tax and Subsidy Level",
xlab = "EU Member",
trace.label = "Fuel Dependence",
col = c("#0198f9", "#f95801"),
lyt = 1,
lwd = 3
)
```

# 4) CONCLUSION 

# Economic Explanations of Fuel Tax and Subsidy Levels
  - It Is Not Surpising that Fossil Fuel Wealth Indicators are Overall Negatively Correlated With Fuel and Tax Levels. 
  
  - Positive relationship between government debt and taxes not surprising. 
  
  - Somehwat positive U shaped relation between GNI per capita and taxes also not            too surprising. 
  
# Political Explanations of Fuel Tax and Subsidy Levels
  -  Slightly positive relation between government effectiveness and tax and subsidy levels also not too surprising
  
  - Idea that Autocracies tax less than democracies also not surprising
  
  - Overall, finding of original paper that political predictors are not strongly correlated with tax and subsidy levels, relative to economic factors, is shown here also. 
  
# There appears to be no interaction between EU member oil dependency on tax/subsidy levels in 2015


# Significance Levels
  - Caveat that only GNI per capita, avetage fuel exports and government debt                coefficients p-values were <0.05
  
  - Other variables greater than 0.05
  
  
# Room for Further Work 
- Analyse the second dataset and look at tax and subsidy changes between 2003-15
  
  
  